---
title: "1. Data transform"
output: html_document
date: "2023-06-09"
---

```{r pack}
r <- getOption("repos")
r["CRAN"] <- "https://cran.biotools.fr/"
options(repos = r)
install.packages("pacman")
library("pacman")
p_load(dplyr)
p_load(rio)
p_load(here)
```

```{r load_data}
modified <- import(here ("data","source","maindata.csv")) # go find the file in the working directory /data/created...this is reproducible should work on all machines if you open the project using the R project file, which sets the wd in the project folder

```

# Data manipulation
## Partner sex column
```{r newvar}

# Add partner_sex column
modified$partner_sex <- ifelse(modified$sexuality == "Hétérosexuel",
                               ifelse(modified$sex == "Homme", "woman", "man"),
                               ifelse(modified$sexuality == "Homosexuel",
                                      ifelse(modified$sex == "Homme", "man", "woman"),
                                      ifelse(modified$sexuality == "Asexuel", "none",
                                             ifelse(modified$sexuality %in% c("Bisexuel", "Pansexuel"), "both", NA))))

```
## Add order column

```{r order}
modified$order <- ifelse(modified$treatment %in% c(4, 5, 6), 2,
                         ifelse(modified$treatment %in% c(1, 2, 3), 1, NA))
```
## Treatment

originally we had 6 treatments, 3 different ones and each could be in 2 possible orders. Now instead we want to combine these into 1 set of treatments.
```{r treat}

modified$treatment2 <- ifelse(modified$treatment == 4, 1,
                              ifelse(modified$treatment == 5, 2,
                                     ifelse(modified$treatment == 6, 3, modified$treatment)))
```
## contraception type

Split type of contraception used now into four separate columns so it can be identified individually.
```{r cont}
split_typenow <- strsplit(as.character(modified$typenow), ",")
modified$first_contnow <- sapply(split_typenow, function(x) ifelse(length(x) >= 1, x[1], NA))
modified$second_contnow <- sapply(split_typenow, function(x) ifelse(length(x) >= 2, x[2], NA))
modified$third_contnow <- sapply(split_typenow, function(x) ifelse(length(x) >= 3, x[3], NA))
modified$fourth_contnow <- sapply(split_typenow, function(x) ifelse(length(x) >= 4, x[4], NA))

# Split typepast column into four separate columns
split_typepast <- strsplit(as.character(modified$typepast), ",")
modified$first_contpast <- sapply(split_typepast, function(x) ifelse(length(x) >= 1, x[1], NA))
modified$second_contpast <- sapply(split_typepast, function(x) ifelse(length(x) >= 2, x[2], NA))
modified$third_contpast <- sapply(split_typepast, function(x) ifelse(length(x) >= 3, x[3], NA))
modified$fourth_contpast <- sapply(split_typepast, function(x) ifelse(length(x) >= 4, x[4], NA))

# Split type_partner column into four separate columns
split_typepartner <-strsplit(as.character(modified$type_partner), ",")
modified$first_contpart <- sapply(split_typepartner, function(x) ifelse(length(x) >= 1, x[1], NA))
modified$second_contpart <- sapply(split_typepartner, function(x) ifelse(length(x) >= 2, x[2], NA))
modified$third_contpart <- sapply(split_typepartner, function(x) ifelse(length(x) >= 3, x[3], NA))
modified$fourth_contpart <- sapply(split_typepartner, function(x) ifelse(length(x) >= 4, x[4], NA))
```
## Location
``` {r loc}
modified$location <- ifelse(modified$date %in% c("2023-05-24", "2023-05-17"), 1, 0)
```
## Age
Create new column for age
```{r age}
# Convert 'date' column to Date format
modified$date <- as.Date(modified$date)

# Extract year and month from 'date'
collected_year <- format(modified$date, "%Y")
collected_month <- format(modified$date, "%m")

# Calculate age based on 'birth_year', 'birth_month', and collected year/month
modified$birth_year <- as.integer(modified$birth_year)
modified$birth_month <- as.integer(modified$birth_month)

age_decimal <- as.numeric(as.Date(paste(collected_year, collected_month, "01", sep = "-"), "%Y-%m-%d") - 
                            as.Date(paste(modified$birth_year, modified$birth_month, "01", sep = "-"), "%Y-%m-%d")) / 365.25

modified$age <- round(age_decimal, 1)
```
## Time writing and time reading
```{r time}

# Convert timestamp columns to POSIXct format
modified$timesubmits1 <- as.POSIXct(modified$timesubmits1, format = "%H:%M:%S")
modified$timesubmits2 <- as.POSIXct(modified$timesubmits2, format = "%H:%M:%S")
modified$timesubmits3 <- as.POSIXct(modified$timesubmits3, format = "%H:%M:%S")

# Calculate timereading as the difference between timesubmits2 and timesubmits1 in seconds
modified$timereading <- as.numeric(difftime(modified$timesubmits2, modified$timesubmits1, units = "secs"))

# Calculate timewriting as the difference between timesubmits3 and timesubmits2 in seconds
modified$timewriting <- as.numeric(difftime(modified$timesubmits3, modified$timesubmits2, units = "secs"))
```
## Contraception classification
Create a simpler classification for contraception use, just whether it is condom or other types of contraception.
``` {r class}

# Create 'contranow' column 
modified$contranow <- ifelse(modified$usecontranow == 'Non', 'none',
                             ifelse(modified$first_contnow == 'Préservatif' &
                                      is.na(modified$second_contnow) &
                                      is.na(modified$third_contnow) &
                                      is.na(modified$fourth_contnow),
                                    'condom', 'contra'))

# Create 'contrapast' column
modified$contrapast <- ifelse(modified$usecontrapast == 'Non', 'none',
                              ifelse(modified$first_contpast == 'Préservatif' &
                                       is.na(modified$second_contpast) &
                                       is.na(modified$third_contpast) &
                                       is.na(modified$fourth_contpast),
                                     'condom', 'contra'))
# Create 'contrapart' column
modified$contrapart <- ifelse(modified$partner == 'Non', 'none',
                              ifelse(modified$first_contpart == 'Préservatif' &
                                       is.na(modified$second_contpart) &
                                       is.na(modified$third_contpart) &
                                       is.na(modified$fourth_contpart),
                                     'condom', 'contra'))
# Create 'contra_overall' column
modified$contra_overall <- ifelse(modified$contranow == 'contra' | modified$contrapast == 'contra', 'contra',
                                  ifelse(modified$contranow == 'condom' | modified$contrapast == 'condom', 'condom',
                                         ifelse(is.na(modified$contranow) & is.na(modified$contrapast), 'none', 'none')))
```

## Male contra type check
Check for instances when men say that they use female-based contraception (now or in the past)

``` {r male hormonal contra check.options}
# Create 'has_other_contra' column for male participants
modified$male_contra_use_now <- ifelse(modified$sex == 'Homme',
                                    ifelse((modified$first_contnow != 'Préservatif' & modified$first_contnow != 'NA' & modified$first_contnow != 'NULL') |
                                           (modified$second_contnow != 'Préservatif' & modified$second_contnow != 'NA' & modified$second_contnow != 'NULL') |
                                           (modified$third_contnow != 'Préservatif' & modified$third_contnow != 'NA' & modified$third_contnow != 'NULL') |
                                           (modified$fourth_contnow != 'Préservatif' & modified$fourth_contnow != 'NA' & modified$fourth_contnow != 'NULL'),
                                           TRUE, FALSE),
                                    FALSE)
# Print the result
summary_table <- table(modified$male_contra_use_now)
print(summary_table)

#two men recorded that they used non-condom contra now.
```

``` {r check male contra use past}

# Create 'has_other_contra' column for contpast columns
modified$male_contra_use_past <- ifelse(modified$sex == 'Homme',
                                            ifelse((modified$first_contpast != 'Préservatif' & modified$first_contpast != 'NA' & modified$first_contpast != 'NULL') |
                                                   (modified$second_contpast != 'Préservatif' & modified$second_contpast != 'NA' & modified$second_contpast != 'NULL') |
                                                   (modified$third_contpast != 'Préservatif' & modified$third_contpast != 'NA' & modified$third_contpast != 'NULL') |
                                                   (modified$fourth_contpast != 'Préservatif' & modified$fourth_contpast != 'NA' & modified$fourth_contpast != 'NULL'),
                                                   TRUE, FALSE),
                                            FALSE)

stpast <- table(modified$male_contra_use_past)
print(stpast) 

# no men said they used female based contraception 
```
```{r edit male contra class}
# for the two individuals identified, change the values

# Update the values of contranow and contra_overall to 'condom' for session_id '7808_23137'
modified$contranow[modified$sessionid == '7808_23137'] <- 'condom'
modified$contra_overall[modified$sessionid == '7808_23137'] <- 'condom'


##### for the person who recorded 'other' in their current type of contraceptive use

modified$contranow[modified$sessionid == '7776_23144'] <- 'none'
modified$contra_overall[modified$sessionid == '7776_23144'] <- 'condom'

```

## Sexuality
Create a simpler column for sexuality
``` {r sexuality}

# Create 'sexualitysum' column
modified$sexualitysum <- ifelse(modified$sexuality == 'Hétérosexuel', 'hetero', 'non-hetero')
```
## Kids
Column for whether participants have children
```{r kids}
modified$kids <- ifelse(modified$child == 'Non', 'no',
                        ifelse(modified$child == 'j\'ai des enfants', 'have', 'yes'))
```
## side-effects
Simpler column for whether participants have experienced side-effects

```{r side}
# Create new column 'sideeffectssim'
modified <- modified %>%
   mutate(sideeffectsim = case_when(
    sideffects %in% c("", "Non", "NA", "NULL") ~ "no",
    sideffects %in% c("Oui, actuellement", "Oui, dans le passé", "Oui, dans le passé,Oui, actuallement") ~ "yes",
    TRUE ~ "yes"
  ))
```
## add new column for partner sex

```{r partner sex}
# Add a new column 'partnersex2' based on 'partner_sex'
modified$partnersex2 <- ifelse(modified$partner_sex %in% c("woman", "both"), "woman", "other")
```

## add column for whether participants use contraception now
```{r contranow}
modified$usecontra <- ifelse(modified$usecontranow == "Oui" | modified$usecontrapast == "Oui", "yes", "no")
```

## add column that classifies subject
``` {r sub class}
modified$Subject <- ifelse(modified$subject == "Médecine" | modified$subject == "Sciences", "Science",
                            ifelse(modified$subject == "Arts" | modified$subject == "Autre" | modified$subject == "Sciences humaines", "Other",
                                   ifelse(modified$subject == "", "Not student", modified$subject)))
```
## Education change column

``` {r edu change}
modified <- modified %>%
  mutate(education_simple = case_when(
    education_level %in% c("Bac +1", "Bac +2", "Bac +3", "Bac +4", "Bac +5") ~ "higher",
    education_level %in% c(
      "Sans diplome", "Certificat d-etude primaire (CEP)", "Troisieme",
      "CAP/BEP/BEPC/BEPS/Brevet élémentaire",
      "Brevet de Technicien, Brevet prof, BEI, BEC, BEA",
      "Baccalaureat technologique ou professionnel",
      "Bac general"
    ) ~ "lower",
    education_level == "" ~ "Not Student",  # If education_level is blank, set as "Not Student"
    TRUE ~ NA_character_  # If education_level doesn't match any condition, set as NA
  ))

```
## rename some variables
``` {r rename}
modified <- modified %>%
  mutate(student = ifelse(student == "", NA, ifelse(student == "Oui", "Yes", ifelse(student == "Non", "No", student))))
modified <- modified %>%
  mutate(location = ifelse(location == "0", "University", ifelse(location == "1", "Zoo", location)))

modified <- modified %>%
  mutate(sex = ifelse(sex == "Homme", "Male", ifelse(sex == "Femme", "Female", sex)))

names(modified)[names(modified) == "sex"] <- "Sex"
names(modified)[names(modified) == "student"] <- "Student"

```

## test
Add a column for the trials that were a test
```{r test}
# Add a column for the 'test' sessions
session_ids <- c(  "679_23145", "9299_23131", "4713_23131", "7662_23131", "427_23131", "7493_23131", "3205_23132", "1457_23135", "7355_23135", "1425_23135", "8763_23135", "5036_23135", "325_23135", "463_23135", "7179_23135",
"343_23137", "343_23143", "1496_23144",               "9299_23131", "151_23131", "4713_23131", "4956_23131", "1866_23131", "5636_23131", "7662_23131", "427_23131",  "7516_23131", "4981_23131", "8133_23131", "4520_23131", "7493_23131", "9698_23131", "4244_23131", "6298_23131",
"981_23131", "8620_23131", "3205_23132", "4705_23134", "9561_23135", "8792_23135", "5928_23135", "9517_23135",
"9241_23135", "8285_23135", "9057_23135", "4400_23135", "4867_23135", "1719_23135", "8942_23135", "8670_23135",
"3173_23135", "9546_23135", "5258_23135", "3738_23135", "3326_23135", "9223_23135", "1457_23135", "7355_23135",
"1135_23135", "6640_23135", "1425_23135", "8763_23135", "6853_23135", "5036_23135", "325_23135", "463_23135",
"6701_23135", "7179_23135", "343_23143", "1496_23144")

# Create the 'test' column
modified$test <- ifelse(modified$sessionid %in% session_ids, "yes", "no")
```
## Data changes
change the records for the data that was wrong
```{r change}
modified[modified$sessionid %in% c("5156_23145", "2875_23144", "4713_23144"), c("treatment", "treatment2")] <- 2
modified[modified$sessionid == "4713_23144", c("birth_year", "age")] <- c(1949, 74.1)

```
# Export

```{r exclusions}

# Save the modified_excluded data frame as a CSV file
export(modified, here ("Data","Created","modified_table.csv")) # go
```
