---
title: "2_scorescombined"
output: html_document
date: "2023-06-09"
---

This script combines scores from Luna + Mattéo + Gabby. Luna is the only one not blind to the experiment.

# Set up
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pack}
r <- getOption("repos")
r["CRAN"] <- "https://cran.biotools.fr/"
options(repos = r)
install.packages("pacman")
library("pacman")
p_load(dplyr)
p_load(rio)
p_load(here)
p_load(readxl)
p_load(irr)
p_load(irrplot)
```

```{r load}
finalscores <- import(here ("Data","Source","finalscores.csv"))
gabyscores <- import(here ("Data","Source","gabyscores.csv"))
```

# Matching scores

For those IDs for which Scorer 1 and Scorer 2 disagree, score from Scorer 3 is taken (as it is the most common of the 3)
```{r load2}
# Define the column names in gabyscores that should be matched
columns_to_match <- c("gmc", "c", "cf", "stom", "blood", "gmt", "alarm", "doors", "gmm", "pain", "fm", "m", "som")

# Update scores in finalscores by matching columns based on the column names
updated_finalscores <- finalscores %>%
  mutate(across(all_of(columns_to_match), ~ ifelse(!is.na(gabyscores[[cur_column()]]) & gabyscores[[cur_column()]] != "", gabyscores[[cur_column()]], .)))

# Output the updated_finalscores data frame
updated_finalscores$cf
export(updated_finalscores, here ("Data","Created","updated_finalscores.csv"))
write.csv(updated_finalscores, file = "updated_finalscores.csv", row.names = FALSE)

# check number of rows
dim(updated_finalscores)
```


# Join scores
```{r join}
modified <- import(here ("data","created","modified_table.csv")) # 363 lines
finalresults <- import(here ("data","created","updated_finalscores.csv")) # 259

finalresults<- finalresults%>%
  select(c("id", "gmc", "c", "cf", "stom", "blood", "gmt", "alarm", "doors", "gmm", "pain", "fm", "m", "som"))


finalresults <- finalresults %>% rename(sessionid = id)
finalresults <- finalresults %>% rename(con = c)
finalresults <- finalresults %>% rename(fertcont = cf)
finalresults <- finalresults %>% rename(fertmed =fm)
finalresults <- finalresults %>% rename(muscle = m)

# Left join modified with finalresults based on sessionid
joined_data <- left_join(modified, finalresults, by = "sessionid")
#check for missing values still
chck<-joined_data %>% select(sessionid, id, fertcont)

# Replace values with NA based on treatment condition (this is because scorers did not not the treatment when scoring and for some of those it does not apply)
# for example, a treatment 1 will get a score of NA for blood, rather than 0.

joined_data$stom <- ifelse(joined_data$treatment2 == 2, NA, joined_data$stom)
joined_data$muscle <- ifelse(joined_data$treatment2 == 2, NA, joined_data$muscle)
joined_data$fertcont <- ifelse(joined_data$treatment2 == 3, NA, joined_data$fertcont)
joined_data$fertmed <- ifelse(joined_data$treatment2 == 3, NA, joined_data$fertmed)
joined_data$som <- ifelse(joined_data$treatment2 == 1, NA, joined_data$som)
joined_data$blood <- ifelse(joined_data$treatment2 == 1, NA, joined_data$blood)
```

# Export
```{r export}
export(joined_data, here ("data","created","scored_data.csv" ))

```

# Measure the agreement between the two main scorers

``` {r import}
sheet1 <- import(here ("Data","Source","lunashortscores.csv"))
sheet2 <- import(here ("Data","Source","matteoshortscores.csv"))
```

```{r extract and reshape}
# Extract the desired columns from both sheets
columns_to_extract <- c("sessionid", "cf", "stom", "blood", "fm", "m", "som")

extracted_sheet1 <- sheet1[, c(1, 5:7, 18:20)]  # Extract columns 5-7 and 18-20
colnames(extracted_sheet1) <- columns_to_extract  # Rename columns

extracted_sheet2 <- sheet2[, c(1, 5:7, 18:20)]  # Extract columns 5-7 and 18-20
colnames(extracted_sheet2) <- columns_to_extract  # Rename columns

columns_to_reshape <- c("cf", "stom", "blood", "fm", "m", "som")

# Reshape the data from wide to long format
sheet1long <- extracted_sheet1 %>%
  select(sessionid, all_of(columns_to_reshape)) %>%
  pivot_longer(cols = -sessionid, names_to = "variable", values_to = "score") %>%
  arrange(sessionid, variable)

sheet2long <- extracted_sheet2 %>%
  select(sessionid, all_of(columns_to_reshape)) %>%
  pivot_longer(cols = -sessionid, names_to = "variable", values_to = "score") %>%
  arrange(sessionid, variable)


# Combine the two sheets into a single data frame
combined_data <- data.frame(sheet1 = sheet1long, sheet2long)

head(combined_data)
```
## Kappa value
At the moment the agreement is tested on all 6 categories used for the analysis. This means that there are 0 scores for 2 text-elements that werent present, depending on the treatment. I've left this in though as the scorers were in fact blind to the treatment.

``` {r measure concordance}
# Calculate the concordance (kappa coefficient) between the two sheets
result <- kappa2(combined_data[, c("sheet1.score", "score")])

#produce a contingency table that shows frequency of agreement / disagreement

contingency_table <- table(combined_data$score, combined_data$sheet1.score)
contingency_table

kappa_result <- kappa2(contingency_table)

# Print the kappa coefficient
print(result$value)


```


