---
title: "3. Data exclude"
output: html_document
date: "2023-06-09"
---

# Set-up
```{r pack}
r <- getOption("repos")
r["CRAN"] <- "https://cran.biotools.fr/"
options(repos = r)
install.packages("pacman")
library("pacman")
p_load(dplyr)
p_load(rio)
p_load(here)
p_load(summarytools)
p_load(DiagrammeR)
p_load(DiagrammeRsvg)
p_load(rsvg)
p_load(DiagrammeR)
p_load(webshot)
p_load(DiagrammeR)
```

```{r load}
excluding <- import(here ("Data","Created","scored_data.csv"))
```

# Exclusions
## Remove test and duplicate rows
```{r exclude}

# Exclude rows where 'test' column is equal to 'yes'
excluding <- excluding[excluding$test != "yes", ]

# remove the duplicate rows from my data
exclude_ids <- c(354, 355, 356, 353, 343, 340, 305, 298, 192, 141, 140, 40, 94, 241, 242)
excluding <- excluding[!(excluding$id %in% exclude_ids), ]

duplicate_rows <- c('1103_23145', '9995_23144')

excluding <- excluding[!(excluding$sessionid %in% duplicate_rows), ]

print(nrow(excluding))

## exclude the participant who asked to be deleted
asked_to_be_deleted_rows <- c('3254_23144')
excluding <- excluding[!(excluding$sessionid %in% asked_to_be_deleted_rows), ]
print(nrow(excluding))
```
## Exclusion justifications
Gradually exclude rows from data set based on different reasons

``` {r exluding}
# Exlcude rows where 'textoutput' is blank OR Null (because participants didn't go that far)

excluding <- excluding[!(is.na(excluding$textoutput) | excluding$textoutput == "" | excluding$textoutput == "NULL"), ]
print(nrow(excluding))

# Specify the rows to be removed based on different reasons
#knew it was a memory test

knew_memory_rows <- c('8309_23132', '7840_23136', '1914_23137', '8982_23137', '1382_23137', '8245_23137','7776_23137', '8645_23137', '118_23137', '4623_23137', '8309_23137', '9948_23137',                '7220_23137', '151_23144', '6311_23144', '1025_23132')

#rushed or didnt fully complete the survey
incomplete_rushed_rows <- c('9257_23136', '6863_23137', '7662_23137', '3961_23144', '5009_23144')

#remaning duplicates
duplicate_rows <- c('1103_23145', '9995_23144')

#French wasnt first language
english_speaker_rows <- c('551_23137')

# Remove rows based on the above different reasons
excluding <- excluding[!(excluding$sessionid %in% incomplete_rushed_rows), ]
print(nrow(excluding))
excluding <- excluding[!(excluding$sessionid %in% knew_memory_rows), ]
print(nrow(excluding))
excluding <- excluding[!(excluding$sessionid =='551_23137'), ]
#participant who didn't take it seriously
excluding <- excluding[!(excluding$sessionid =='6810_23136'), ]
print(nrow(excluding))

# Update specific rows with new values (two cases where the participants had to reload their survey and therefore their treatments were incorrect)
excluding[excluding$sessionid %in% c("5156_23145", "2875_23144", "4713_23144"), c("treatment", "treatment2")] <- 2
#one lady where the birth year didnt go back enough so had to enter manually
excluding[excluding$sessionid == "4713_23144", c("birth_year", "age")] <- c(1949, 74.1)

#exclude the sex =other
excluding <- excluding[excluding$Sex != "Autre", ]

print(nrow(excluding))

```

# Examine exclusions
```{r examin_exclu}

## examine the excluded data

view(dfSummary(excluding))
print(nrow(excluding))
```

# Renaming and selecting out columns
```{r renaming}

############## changing the treatment column names and exluding columns not needed

# Rename columns
excluding <- excluding %>%
  rename(fulltreatment = treatment, treatment = treatment2)

# Exclude columns
excluding <- excluding %>%
  select(-birth_year, -birth_month, -sexeautre, -genreautre, -timesubmits1,
         -timesubmits2, -textinputp1, -textinputp2, -textinputp3,
         -timesubmits3, -timesubmits4, -third_contnow, -fourth_contnow,
         -third_contpart, -fourth_contpart)

# Calculate the median and SD

median_age <- median(excluding$age)
sd_age <- sd(excluding$age)

# Output the median and SD
print(paste("Median:", median_age))
print(paste("SD:", sd_age))

# Calculate quartiles
age_quartiles <- quantile(excluding$age, probs = c(0, 0.25, 0.5, 0.75, 1))
# Assign new column of age categories based on quartiles
excluding$agecat <- cut(excluding$age, breaks = age_quartiles, labels = c("Q1", "Q2", "Q3", "Q4"), include.lowest = TRUE)

#
print(nrow(excluding))
```
# Renaming variables
``` {r rename}
colnames(excluding) <- c("id",	"Sessionid",	"Fulltreatment",	"date",	"Student",	"university",	"subjectbroad",	"gender",	"Sex",	"parent_degree",	"education_level",	"textoutput",	"contrausenow",	"typecontranow",	"contrausepast",	"typecontrapast",	"plannedcontrachange",	"partnercontrause",
"typecontrapartner",	"sideffects",	"stopcozside",	"sevside",	"sexuality",	"sexualityother",	"child",	"timewriting",	"partner_sex",	"order",	"treatment",	"first_contranow",	"second_contranow",	"first_contrapast",	"second_contrapast",
"third_contrapast",	"fourth_contrapast",	"first_contrapart",	"second_contrapart",	"age",	"timereading",	"contranowsimple",	"contrapastsimple",	"contrapartsimple",	"contra_overall", "currmalecont", "pastmalecont",	"sexualitysimple",	"kidsimple",	"sideffectssim",	"partnersexsimple",	"usecontraever",	"Subject", "education_simple", "location",	"test",	"gmc",	"con",	"fertcont",	"stom",	"blood",	"gmt",	"alarm",	"doors",	"gmm",	"pain",	"fertmed",	"muscle",	"som",	"agecat")


```

# Export
```{r export}
export(excluding,here ("Data","Created","data_250.csv"))

```

#Diagram of exclusion process

```{r , echo=FALSE}

grViz("
  digraph {
  layout = dot
    node [shape = rectangle,color=LIGHTBLUE,style=filled,fontsize=45,fixedsize=FALSE,width=4.0]
    edge[color=grey,arrowhead=vee,penwidth=5,arrowsize=5]
    A[label = '289 consenting participants']
    
    C[label = '262 complete survey responses']
    
    E[label = '252 participants adhering to study conditions']
    
    G[label = '251 participant with French as a main language: n=251']
    
    I[label = '250 participants who provided information on their sex']
    
    
    node [shape = rectangle,color=GRAY,style=filled,fontsize=45,fixedsize=FALSE,width=4.0]
    edge[color=grey,arrowhead=vee,penwidth=5,arrowsize=5]
    B[label = '27 participants excluded for not properly completing the survey or writing nothing']
    D[label = '11 Participants excluded for understanding study purpose']
    F[label = '1 participant excluded for not speaking French as a main language']
    H[label = '1 participant excluded for recording their sex as neither male or female']
    
    
    node [shape = rectange,color=orange,style=filled,fontsize=45,fixedsize=FALSE,width=4.0]
    edge[color=grey,arrowhead=vee,penwidth=5,arrowsize=4]
    tab6 [label = 'A: Fertility vs Health: n=83'];
    tab7 [label = 'B: Fertility vs Neutral: n=84'];
    tab8 [label = 'C: Health vs Neutral: n=83'];
  
blank1[label = '', width = 0.01, height = 0.01]   
A -> blank1[dir=none];
blank1 -> B[minlen=10];
  {{ rank = same; blank1 B }}
blank1 -> C

blank2[label = '', width = 0.01, height = 0.01]   
C -> blank2[dir=none];
blank2 -> D[minlen=10];
  {{ rank = same; blank2 D }}
blank2 -> E

blank3[label = '', width = 0.01, height = 0.01]   
E -> blank3[dir=none];
blank3 -> F[minlen=10];
  {{ rank = same; blank3 F }}
blank3 -> G

blank4[label = '', width = 0.01, height = 0.01]   
G -> blank4[dir=none];
blank4 -> H[minlen=10];
  {{ rank = same; blank4 H }}
blank4 -> I

I -> tab6 [minlen=4];
I -> tab7 [minlen=4];
I -> tab8 [minlen=4];

    
  }
   ")



```

  tab1.1 [label = '26 participants excluded for not properly completing the survey or writing nothing'];
      tab1.2 [label = '11 participants excluded for understanding study purpose'];
      tab1.3 [label = '1 participant excluded for not speaking French as a main language'];
      tab1.4 [label = '1 participant excluded for not recording their sex as male or female'];
    }
    
    'Excluding participants who did not complete the entire survey, \n had to rush or wrote nothing: n=263'];
    tab3 [label = 'Excluding participants who knew it was a memory test \n or about contraception (e.g. from observing other participants) : n=252'];
    tab4 [label = 'Excluding participant who did not speak French as a first language: n=251'];
    tab5 [label = 'Excluding participant whose sex was neither male nor female: n=250'];
```


``` {r flow diagram}

















# Create the chart with the 'additional block' flowing from the first block to the right
chart <- grViz("
  digraph flowchart {
    graph [overlap = true, fontsize = 24, rankdir = LR];  # Adjusted rankdir to LR (left-to-right layout)
    
    # Main blocks
    node [fontname = Arial, shape = rectangle, style = filled, fillcolor = grey, fontcolor = black];
    tab1 [label = 'Study completed by consenting participants: n=289'];
    tab2 [label = 'Excluding participants who did not complete the entire survey, \n had to rush or wrote nothing: n=263'];
    tab3 [label = 'Excluding participants who knew it was a memory test \n or about contraception (e.g. from observing other participants) : n=252'];
    tab4 [label = 'Excluding participant who did not speak French as a first language: n=251'];
    tab5 [label = 'Excluding participant whose sex was neither male nor female: n=250'];

    # Additional block
    node [shape = rectangle, fillcolor = white, fontcolor = black];
    newblock [label = 'Additional Block'];
  
    # Main block connections
    tab1 -> tab2 -> tab3 -> tab4 -> tab5
  
    # Connection from tab1 to newblock
    rank=same tab1 -> newblock;
  }
")

chart <- grViz("
  digraph flowchart {
    graph [overlap = true, fontsize = 24];
    
    node [fontname = Arial, shape = rectangle, style = filled, fillcolor = grey, fontcolor = black];
    tab1 [label = 'Study completed by consenting participants: n=289'];
    tab2 [label = 'Excluding participants who did not complete the entire survey, \n had to rush or wrote nothing: n=263'];
    tab3 [label = 'Excluding participants who knew it was a memory test \n or about contraception (e.g. from observing other participants) : n=252'];
    tab4 [label = 'Excluding participant who did not speak French as a first language: n=251'];
    tab5 [label = 'Excluding participant whose sex was neither male nor female: n=250'];
    
    node [shape = rectangle, fillcolor = white, fontcolor = black];
    tab6 [label = 'Treatment 1: Fertility vs Health: n=83'];
    tab7 [label = 'Treatment 2: Fertility vs Neutral: n=85'];
    tab8 [label = 'Treatment 3: Health vs Neutral: n=83'];
    
      # Additional block
    node [shape = rectangle, fillcolor = white, fontcolor = black];
    newblock [label = 'Additional Block'];
  
  
    tab1 -> tab2 -> tab3 -> tab4 ->tab5;
    tab5 -> tab6;
    tab5 -> tab7;
    tab5 -> tab8;
    
       # Connection from tab1 to newblock
    tab1 -> newblock [ltail = mainblocks, lhead = newblock];
  
  }
")

chart <- grViz("
  digraph flowchart {
    graph [overlap = true, fontsize = 24, rankdir = TB];  # Adjusted rankdir to TB (top-to-bottom layout)
    
    # Main blocks
    node [fontname = Arial, shape = rectangle, style = filled, fillcolor = grey, fontcolor = black];
    subgraph mainblocks {
      tab1 [label = 'Study completed by consenting participants: n=289'];
      tab2 [label = 'Excluding participants who did not complete the entire survey, \n had to rush or wrote nothing: n=263'];
      tab3 [label = 'Excluding participants who knew it was a memory test \n or about contraception (e.g. from observing other participants) : n=252'];
      tab4 [label = 'Excluding participant who did not speak French as a first language: n=251'];
      tab5 [label = 'Excluding participant whose sex was neither male nor female: n=250'];
    }
    
    # Side bubbles
    node [shape = rectangle, fillcolor = white, fontcolor = black];
    subgraph sidebubbles {
      tab1.1 [label = '26 participants excluded for not properly completing the survey or writing nothing'];
      tab1.2 [label = '11 participants excluded for understanding study purpose'];
      tab1.3 [label = '1 participant excluded for not speaking French as a main language'];
      tab1.4 [label = '1 participant excluded for not recording their sex as male or female'];
    }
  
    # Main block connections
    tab1 -> tab2 -> tab3 -> tab4 -> tab5;
    
    # Side bubble connections
    tab1 -> tab1.1 [lhead = sidebubbles];
    tab2 -> tab1.2 [lhead = sidebubbles];
    tab3 -> tab1.3 [lhead = sidebubbles];
    tab4 -> tab1.4 [lhead = sidebubbles];
  }
")
# Display chart on the console
print(chart)

# Save as PDF
webshot(grVizOutput(chart), file = "Figure2.pdf", cliprect = "viewport", vwidth = 1000, vheight = 800)

# Save as PNG
webshot(grVizOutput(chart), file = "Figure2.png", cliprect = "viewport", vwidth = 1000, vheight = 800)
```
