---
title: "5_logistic regreesion"
output: html_document
date: "2023-07-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Load packages
These packages werent working with p_load so have used the traditional way.

```{r pack, include=FALSE}
r <- getOption("repos")
r["CRAN"] <- "https://cran.biotools.fr/"
options(repos = r)
install.packages("pacman")
library("pacman")
library(pacman)
install.packages("tidyverse")
library(tidyverse)
install.packages("rio")
library(rio)
install.packages("rlang")
library(rlang)
install.packages("here")
library(here)
install.packages("summarytools")
library(summarytools)
install.packages("dplyr")
library(dplyr)
install.packages("gtsummary")
library(gtsummary)
install.packages("car")
library(car)
install.packages("lme4")
library(lme4)
install.packages("ggplot2")
library(ggplot2)
install.packages("gt")
library(gt)
install.packages("glmmTMB")
library(glmmTMB)
install.packages(blme)
library(blme)
install.packages(sjPlot)
library(sjPlot)
install.packages("effects")
library(effects)
install.packages("DHARMa")
library(DHARMa)
```

#### Reset repro (if needed)
If the packages stop working then resetting the repository might help.

```{r repro}
options(repos = getOption("repos")["CRAN"])
```

## Load data
load data that has both the scores and metadata
```{r load_data, include=FALSE}
wide_data<-import(here ("Data","Created","data_250.csv"))
print(ncol(wide_data))
head(wide_data)
```
# Examine
```{r exam1}
view(dfSummary(wide_data))
```

## Exclude unnecessary columns
```{r streamline}

columns_to_exclude <- c(
  "Fulltreatment", "date", "gender", "education_level", "textoutput", "typecontranow", "subjectbroad",
  "typecontrapast", "sexuality", "sexualityother", "child", "order", "first_contranow",
  "second_contranow", "first_contrapast", "second_contrapast", "third_contrapast",
  "fourth_contrapast", "first_contrapart", "second_contrapart", "test", "typecontrapartner", "sideffects", "partner_sex"
)

# Exclude the specified columns
wide_data<- wide_data[, !(names(wide_data) %in% columns_to_exclude)]

# Print the filtered data frame
print(ncol(wide_data))
print(names(wide_data))
```

## Change format
What previously was neutral is now HealthMild and Health is HealthModerate

```{r def}
# Define the selected columns including text elements and meta data
selected_columns <- c("Sessionid", "fertcont", "stom", "blood", "fertmed", "muscle","som", "alarm", "doors")

# Define a function to assign the context based on the text element
assign_context <- function(text_element) {
  Context <- case_when(
    text_element %in% c("fertcont", "stom", "blood") ~ "Contraception",
    text_element %in% c("alarm", "doors") ~ "tech",
    text_element %in% c("fertmed", "muscle", "som") ~ "Medical",
    TRUE ~ NA_character_
  )
  return(Context)
}

## convert data from wide to long format and add the context column

long_data <- wide_data %>%
  select(all_of(selected_columns)) %>%
  pivot_longer(cols = -Sessionid,
               names_to = "text_element",
               values_to = "score") %>%
  mutate(Context = assign_context(text_element))

# Perform a left join with the original wide-format data
fld <- left_join(long_data, wide_data, by = "Sessionid")

```

## Run this code to treat the health side effects as two separate categories, moderate and mild. 
```{r normal analysis option}

# Create the Side_effect column based on text_element values
fld$"Side_effect" <- ifelse(fld$text_element %in% c("fertcont", "fertmed"), "Fertility", ifelse(fld$text_element %in% c("muscle", "stom"), "HealthModerate",  ifelse(fld$text_element %in% c("som", "blood"), "HealthMild", NA)))
```

## OPTIONAL  - Run this code instead of chunk above to run combine the two health side effects together, this gives worse AIC values for the main models. 
``` {r OPTIONAL change analysis}

### try grouping all health together
# Create the Side_effect column based on text_element values
fld$"Side_effect" <- ifelse(fld$text_element %in% c("fertcont", "fertmed"), "Fertility", ifelse(fld$text_element %in% c("muscle", "stom", "som", "blood"), "Health", NA))
```

``` {r work}
#remove the rows that didn't have a score, because the treatment didn't include that text element and filter out the technology context

fld<-fld %>%
  filter(score!="NA")%>%
  filter(Context!="tech")

#rename all data
alldata<-fld
```
# Examine
```{r exam}
view(dfSummary(alldata))
```

# Relevel and set factors
## Run this chuck for treating health side-effects separately
```{r releve}
alldata$Side_effect <- as.factor(alldata$Side_effect)
alldata$Side_effect<- relevel(alldata$Side_effect, ref = "HealthMild")
```

## Run this chunk instead of the one above to group the health side-effects together
``` {r relevle}

alldata$Side_effect <- as.factor(alldata$Side_effect)
alldata$Side_effect<- relevel(alldata$Side_effect, ref = "Health")
```

```{r full relevel}

alldata$treatment<- factor(alldata$treatment)
alldata$treatment<- relevel(alldata$treatment, ref = "3")

alldata$Context<- factor(alldata$Context)
alldata$Context<- relevel(alldata$Context, ref = "Medical")

alldata$Sex <- factor(alldata$Sex)
alldata$Sex<- relevel(alldata$Sex, ref = "Female")

alldata$Student <- factor(alldata$Student)
alldata$Student<- relevel(alldata$Student, ref = "No")

alldata$university <- factor(alldata$university)
alldata$university <- relevel(alldata$university, ref = "Université de Montpellier")

alldata$Sessionid <- factor(alldata$Sessionid)

alldata$parent_degree<-factor(alldata$parent_degree)
alldata$parent_degree<- relevel(alldata$parent_degree, ref = "Aucun")

alldata$contrausenow<-factor(alldata$contrausenow)

alldata$plannedcontrachange<- factor(alldata$plannedcontrachange)

alldata$partnercontrause<-factor(alldata$partnercontrause)
alldata$partnercontrause<- relevel(alldata$partnercontrause, ref = "Non")

alldata$stopcozside<-factor(alldata$stopcozside)
alldata$stopcozside<-relevel(alldata$stopcozside, ref = "Non")

alldata$sevside<-factor(alldata$sevside)
alldata$sevside<-relevel(alldata$sevside, ref = "Léger")

alldata$location<-factor(alldata$location)

alldata$contranowsimple<-factor(alldata$contranowsimple)
alldata$contranowsimple<-relevel(alldata$contranowsimple, ref = "condom")

alldata$contrapastsimple<-factor(alldata$contrapastsimple)
alldata$contrapastsimple<-relevel(alldata$contrapastsimple, ref = "condom")

alldata$contrapartsimple<-factor(alldata$contrapartsimple)
alldata$contrapartsimple<-relevel(alldata$contrapartsimple, ref = "condom")

alldata$contra_overall<-factor(alldata$contra_overall)
alldata$contra_overall<-relevel(alldata$contra_overall, ref = "condom")

alldata$plannedcontrachange<-factor(alldata$plannedcontrachange)
alldata$plannedcontrachange<-relevel(alldata$plannedcontrachange, ref = "Non")

alldata$sexualitysimple<-factor(alldata$sexualitysimple)

alldata$kidsimple<-factor(alldata$kidsimple)
alldata$kidsimple<-relevel(alldata$kidsimple, ref = "no")

alldata$sideffectssim<-factor(alldata$sideffectssim)

alldata$partnersexsimple<-factor(alldata$partnersexsimple)

alldata$usecontraever<-factor(alldata$usecontraever)

alldata$Subject<-factor(alldata$Subject)
alldata$Subject<- relevel(alldata$Subject, ref = "Other")

alldata$education_simple<-factor(alldata$education_simple)

```

## Distribution of variables - main data

```{r summary}
library(dplyr)

#need to rename one variable for it to appear

wide_data$sideffectssim <- ifelse(wide_data$sideffectssim == 'no', 'Non', 'Oui')

#choose the variables you want in the table, then rename them in word.

maintable <- wide_data[, c("treatment", "agecat", "Subject", "Sex", "partnersexsimple", "contra_overall", "sideffectssim")]

tbl_summary(maintable, by = treatment) %>% 
  add_p() %>%
  add_overall %>%
 add_stat_label() #%>%
gt::tab_options(table.font.names = "Times New Roman")
```

## Distribution of main variables - for annex that includes all variables

```{r distrib for annex}
## Distribution of variables - main data

library(dplyr)
#choose the variables you want in the table, then rename them in word.

#need to rename variables to appear in table

wide_data$usecontraever <- ifelse(wide_data$usecontraever == 'no', 'Non', 'Oui')

wide_data$Student <- ifelse(wide_data$Student == 'No', 'Non', 'Oui')

wide_data$sideffectssim <- ifelse(wide_data$sideffectssim == 'no', 'Non', 'Oui')


annextable <- wide_data[, c("treatment", "agecat", "Subject", "Sex", "partnersexsimple", "contra_overall", "sideffectssim", "Student", "university", "parent_degree", "contrausenow", "contrausepast", "plannedcontrachange", "partnercontrause", "stopcozside", "sevside", "location", "age", "contranowsimple", "contrapastsimple", "contrapartsimple", "contra_overall", "sexualitysimple", "kidsimple", "usecontraever", "education_simple")]

tbl_summary(annextable, by = treatment) %>% 
  add_p() %>%
  add_overall %>%
 add_stat_label() #%>%
gt::tab_options(table.font.names = "Times New Roman")

```

# Univariate analysis
```{r univar anal}
#0. treatment
model0 <- glmer(score ~ treatment + (1 | Sessionid), data = alldata, family = binomial)

#1. Side_effect
model1 <- glmer(score ~ `Side_effect` + (1 | Sessionid), data = alldata, family = binomial)

#2. Context
model2 <- glmer(score ~ Context + (1 | Sessionid), data = alldata, family = binomial)

#3. log Age - signif
model3 <- glmer(score ~ log(age) +(1 | Sessionid), data = alldata, family = binomial)

#4. age cat - non sig
model4 <- glmer(score ~  agecat + (1 | Sessionid), data = alldata, family = binomial)

#5. sex - using data excluding other - slightly sig - MAIN
model5 <- glmer(score ~ Sex + (1 | Sessionid), data = alldata, family = binomial)

#6. student status - sig
model6 <- glmer(score ~  Student + (1 | Sessionid), data = alldata, family = binomial)

#7. university
model7 <- glmer(score ~  university + (1 | Sessionid), data = alldata, family = binomial)

#8 Subject (broad terms)
model8 <- glmer(score ~ Subject + (1 | Sessionid), data = alldata, family = binomial)

#9. location - sig -MAIN
model9 <- glmer(score ~  location + (1 | Sessionid), data = alldata, family = binomial)

#10. parent degree - not sig
model10 <- glmer(score ~  parent_degree+ (1 | Sessionid), data = alldata, family = binomial)

#11. education level
model11 <- glmer(score ~ education_simple + (1 | Sessionid), data = alldata, family = binomial)

#12. use contra now - not sig
model12 <- glmer(score ~  contrausenow+ (1 | Sessionid), data = alldata, family = binomial)

#13. use contra past - not sig, but below 0.2 - ANNEX
model13 <- glmer(score ~  contrausepast+ (1 | Sessionid), data = alldata, family = binomial)

#14. Partner contra use
model14 <- glmer(score ~  partnercontrause + (1 | Sessionid), data = alldata, family = binomial)

#15. simple contra type used now
model15 <- glmer(score ~  contranowsimple + (1 | Sessionid), data = alldata, family = binomial)

#16. type contra used past
model16 <- glmer(score ~  contrapastsimple + (1 | Sessionid), data = alldata, family = binomial)

#17. type contra used partner
model17 <- glmer(score ~  contrapartsimple + (1 | Sessionid), data = alldata, family = binomial)

#18. Type contra used overall (present or past)
model18 <- glmer(score ~  contra_overall + (1 | Sessionid), data = alldata, family = binomial)

#19. use contra ever (used contra present or past)
model19 <- glmer(score ~ usecontraever   + (1 | Sessionid), data = alldata, family = binomial)

#20. use contra past - not sig, but below 0.2 - ANNEX
### SUBSET??
model20 <- glmer(score ~  plannedcontrachange + (1 | Sessionid), data = alldata, family = binomial)

#21. simple side effects - not sig - ANNEX
### SUBSET?
model21 <- glmer(score ~ sideffectssim + (1 | Sessionid), data = alldata, family = binomial)

#22 stop because side
model22 <- glmer(score ~ stopcozside + (1 | Sessionid), data = alldata, family = binomial)

#23 severity side
model23 <- glmer(score ~ sevside + (1 | Sessionid), data = alldata, family = binomial)

#24. kids simple (want, don't want, have) - not sig - ANNEX
model24 <- glmer(score ~ kidsimple + (1 | Sessionid), data = alldata, family = binomial)

#25. partner sex simple (woman / other)
model25 <- glmer(score ~ partnersexsimple + (1 | Sessionid), data = alldata, family = binomial)

#26. sexuality simple (hetero / non-hetero)
model26 <- glmer(score ~ sexualitysimple + (1 | Sessionid), data = alldata, family = binomial)

```
# Univariate model summaries 

``` {r m1}
#1. Side_effect
summary(model1)
```

```{r m2}
#2. Context
summary(model2)
```

```{r m3}
#3. log Age - signif
summary(model3)
```

```{r m4}
#4. age cat - non sig
summary(model4)
```

```{r m5}
#5. sex - using data excluding other - slightly sig - MAIN
summary(model5)
```

```{r m6}
#6. gender
summary(model6)
```

```{r m7}
#7. student status - sig
summary(model7)
```

```{r m8}
#8. university
summary(model8)
```

```{r m9}
#9 Subject (broad terms)
summary(model9) 
```

``` {r m10}
#10. location - sig -MAIN
summary(model10)
```

``` {r m11}
#11. parent degree - not sig
summary(model11)
```

```{r m12}
#12. education level
summary(model12)
```

``` {r m13}
#13. use contra now - not sig
summary(model13) 
```

``` {r m14}
#14. use contra past - not sig, but below 0.2 - ANNEX
summary(model14)
```

```{r m15}
#15. Partner contra use
summary(model15)
```

```{r m16}
#16. simple contra type used now
summary(model16) 
```

``` {r m17}
#17. type contra used past
summary(model17) 
```

``` {r m18}
#18. type contra used partner
summary(model18) 
```

```{r m19}
#19. Type contra used overall (present or past)
summary(model19)
```

``` {r m20}
#20. use contra ever (used contra present or past)
summary(model20)
```

```{r m21}
#21. use contra past - not sig - ANNEX
### SUBSET??
summary(model21)
```

``` {r m22}
#22. simple side effects - not sig - ANNEX
### SUBSET?
summary(model22)
```

``` {r m23}
#23 stop because side
summary(model23)
```

``` {r m24}
#24 severity side
summary(model24)
```

``` {r m25}
#25. kids simple (want, don't want, have) - not sig - ANNEX
summary(model25) <- glmer(score ~ kidsimple + (1 | Sessionid), data = alldata, family = binomial)
```

``` {r m26}
#26. partner sex simple (woman / other)
summary(model26)
```

```{r m27}
#27. sexuality simple (hetero / non-hetero)
summary(model27)
```

# Univariate analysis table
guide from this website: https://cran.r-project.org/web/packages/gtsummary/vignettes/tbl_regression.html 
``` {r univar table}

#femaleonly <- alldata[alldata$Sex == "Female", ]

alldata %>%
  
   select(treatment, Sessionid, score, Side_effect, Context, agecat, Sex, 
          Student, university, Subject, location, parent_degree, education_simple, contrausenow, contrausepast, partnercontrause, contranowsimple, contrapastsimple, contrapartsimple, contra_overall, usecontraever, plannedcontrachange, sideffectssim, stopcozside, sevside, kidsimple, partnersexsimple, sexualitysimple) %>%
  
    tbl_uvregression(
        method = glmer,
        y = score,
        method.args = list(family = binomial),
        exponentiate = TRUE,
        pvalue_fun = function(x) style_pvalue(x, digits = 2),
        formula = "{y} ~ {x}+ (1|Sessionid)"
    ) # %>%
 # add_global_p() %>% # add global p-value
#  add_nevent() %>% # add number of events of the outcome
 # add_q() %>% # adjusts global p-values for multiple testing
#  bold_p() %>% # bold p-values under a given threshold (default 0.05)
#  bold_p(t = 0.10, q = TRUE) %>% # now bold q-values under the threshold of 0.10
# bold_labels()

```

# Testing basic and null models

```{r nulls}

modelnull<- glmer(score ~ 1 + (1 | Sessionid), data = subset(alldata), family = binomial(link='logit'))
#summary(modelnull)

modelbasic1 <- glmer(score ~ `Side_effect` + (1 | Sessionid), data = alldata, family = binomial(link='logit'))
#summary(modelbasic1)

modelbasic2 <- glmer(score ~ Context + (1 | Sessionid), data = alldata, family = binomial(link='logit'))
#summary(modelbasic2)

modelbasic3 <- glmer(score ~ `Side_effect` + Context + (1 | Sessionid), data = alldata, family = binomial(link='logit'))
#summary(modelbasic3)
```

## Test models with variables with p-value <0.2

``` {r testing univaraite variables}

#education level - not significant
modelunivar1 <- glmer(score ~ `Side_effect` + Context + Sex + education_simple + (1 | Sessionid), data = alldata, family = binomial(link='logit'))
#summary(modelunivar1)

#past contraceptive use - not significant
modelunivar2 <- glmer(score ~ `Side_effect` + Context + Sex + contrausepast + (1 | Sessionid), data = alldata, family = binomial(link='logit'))
#summary(modelunivar2)

#type of contraception used in past - not significant
modelunivar3 <- glmer(score ~ `Side_effect` + Context + Sex + contrapastsimple + (1 | Sessionid), data = alldata, family = binomial(link='logit'))
#summary(modelunivar3)

#experienced contraceptive Side_effects - not significant
modelunivar4 <- glmer(score ~ `Side_effect` + Context + Sex + sideffectssim + (1 | Sessionid), data = alldata, family = binomial(link='logit'))
#summary(modelunivar4)

#severity of side effects - doesn't converge, even not significant
modelunivar5 <- glmer(score ~ `Side_effect` + Context + Sex + sevside + (1 | Sessionid), data = alldata, family = binomial(link='logit'))
#summary(modelunivar5)

```

# Final models options
```{r testing}
modelfinal1 <- glmer(score ~ `Side_effect` + Context + Sex + (1 | Sessionid), data = alldata, family = binomial(link='logit'))
#summary(modelfinal1)

modelfinal2 <- glmer(score ~ `Side_effect` + Context + Sex + location + (1 | Sessionid), data = alldata, family = binomial(link='logit'))
#summary(modelfinal2)

modelfinal3 <- glmer(score ~ `Side_effect` + Context + Sex + Student + (1 | Sessionid), data = alldata, family = binomial(link='logit'))
#summary(modelfinal3)

modelfinal4 <- glmer(score ~ `Side_effect` + Context + location + Student + (1 | Sessionid), data = alldata, family = binomial(link='logit'))
#summary(modelfinal4)

modelfinal5 <- glmer(score ~ `Side_effect` + Context + Sex + Subject + (1 | Sessionid), data = alldata, family = binomial(link='logit'))
#summary(modelfinal5)


```
## grouped health AIC = 922.8644
## separate health AIC = 921.0666
``` {r model selection}

AIC(modelnull)
AIC(modelbasic1)
AIC(modelbasic2)
AIC(modelbasic3)
AIC(modelfinal1)
AIC(modelfinal2)
AIC(modelfinal3)
AIC(modelfinal4)
AIC(modelfinal5)

###### so models 5 gives lowest AIC values (both within 1 of eachother)



```

# Model selection plot - annex

```{r plot of model selection}
p2<-tab_model(modelfinal2, modelfinal3, modelfinal4, show.aic = T)
p2
```

# Model selection plot - main text

```{r plot of model selection}
p<-tab_model(modelbasic3, modelfinal1, modelfinal5, show.aic = T)
p
```

# Interactions

``` {r interact}
modelinter1 <- glmer(score ~ `Side_effect`*Sex + (1 | Sessionid), data = subset(alldata), family = binomial(link='logit'))
summary(modelinter1)

modelinter2 <- glmer(score ~ Context*Sex + (1 | Sessionid), data = subset(alldata), family = binomial(link='logit'))
summary(modelinter2)

modelinter3 <- glmer(score ~ `Side_effect` * Context + (1 | Sessionid), data = alldata, family = binomial(link="logit"))
summary(modelinter3)

```

# Check model assumptions
good resource: https://towardsdatascience.com/assumptions-of-logistic-regression-clearly-explained-44d85a22b290 
```{r model perf}
##this works when you make the plot window very big
performance::check_model(modelfinal5)
#took the binned residuals plot from this

#test for multicolinearity. VIF value below 5 suggests lack of multicolinearity
car::vif(modelfinal5)

# 2. No complete separation (only applicable to binary logistic regression models)
## i dont think this test is very useful.
check_complete_separation <- function(modelfinal5) {
  predictors <- all.vars(formula(modelfinal5))[-1]
  complete_separation <- sapply(predictors, function(predictor) {
    if (nlevels(alldata[[predictor]]) == 2) {
      tryCatch(vcdExtra::mosaicglm(as.formula(paste0("score ~ ", predictor)), data = alldata), error = function(e) TRUE)
    } else {
      FALSE
    }
  })
  if (any(complete_separation)) {
    cat("Warning: Complete separation detected for the following predictors:\n")
    cat(paste0(" - ", predictors[complete_separation]), "\n")
    cat("Consider removing these predictors causing complete separation.\n")
  }
}

check_complete_separation(modelfinal5)

# 3. Adequate sample size
# Ensure that each category of categorical predictors has a sufficient number of observations

# 4. Proportional odds assumption (only applicable to ordinal categorical predictors)
# Example for 'Context' as an ordinal categorical variable
if ("Context" %in% names(alldata)) {
  ordinal_model <- polr(Context ~ 1, data = alldata, method = "logistic")
  brant_test <- brant(ordinal_model, modelfinal5)
  print(brant_test)
}

# 5. Independence of observations
# Assumes your data is already appropriately randomized or sampled to ensure independence

```

```{r outlier check}
## check for strongly influential outliers
# Get influence measures



```

# Effect plots

``` {r effect plots}
#need to rename column for purpose of effect plot

# Plot effect plot
plot(allEffects(modelfinal5))

```

# Forest plot

``` {r bar}
plot_model(modelfinal5, order.terms = c(1, 2, 3, 4, 5, 6), vline.color = "red", show.values = TRUE, value.offset = .3, title = "")
```

# Random effect strucutre

test nesting the models with different random effects
```{r random effects}
m1 <- glmmTMB(score ~ 1+ (1 | Sessionid), data = alldata, family = binomial)
m2 <- glmmTMB(score ~ 1+ (1 | Sessionid/Context), data = alldata, family = binomial)
m3 <- glmmTMB(score ~ 1+ (1 | Sessionid/location), data = alldata, family = binomial)
m4 <- glmmTMB(score ~ 1 + (1 | location), data = alldata, family = binomial)
AIC(m1)
AIC(m2)
AIC(m3)
AIC(m4)

# AIC is higher for m2 and not so different - most variance due to inter-individual variance
```


